Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(dplyr)
library(ggplot2)
pf <- read.delim('pseudo_facebook.tsv')

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()


pf.fc_by_age_gender <- group_by(pf,age,gender) %>%
  filter(!is.na(gender)) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n=n()) %>%
  ungroup() %>% 
  arrange(age)

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y=median_friend_count), data=pf.fc_by_age_gender) +
  geom_line(aes(color=gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female/male),
      data=pf.fc_by_age_gender.wide) +
   geom_line() +
   geom_hline(aes(yintercept = 1), linetype=2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004,2009,2011,2012,2014))
table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y = mean)+
  geom_line(fun.y = mean, stat='summary', linetype=2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure > 1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure),
       data = subset(pf, tenure>1))+
  geom_line(aes(color=year_joined.bucket))
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
summary(yo)
str(yo)
yo$id <- factor(yo$id)

#qplot(data=yo, x = price, fill=I('#F79420'), binwidth=10)
qplot(data=yo, x=price, fill=I('#F79420'))

ggplot(aes(x=price), data=yo) +
  geom_histogram(stat='bin', binwidth =10)

ggplot(aes(x=price), data=yo) +
  geom_histogram(stat='bin')
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
table(yo$id)
all.purchases <- transform(yo,table(yo$id))
yo <- transform(yo, all.purchases=strawberry+blueberry+pina.colada+plain+mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x=all.purchases), data = yo) +
  geom_histogram(binwidth=1)

ggplot(aes(x=time,y=price), data=yo) +
  geom_point(alpha=1/20)
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(10000)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x=time, y= price), data= subset(yo, id %in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size=all.purchases), pch=1) +
  ggsave('seed_10_000.jpg')
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
install_github("ggobi/ggally")
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
head(pf)
pf_subset <- pf[,c('age', 'friend_count', 'friendships_initiated', 'mobile_likes')]
pf_subset <- pf[,c('gender', 'friendships_initiated', 'friend_count', , 'age',  'mobile_likes')]
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),-1])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

* We have look at many variables at once and plotting them.
* We explore how to compare many variables
* We overcome complexity of our data with sampling and smoothing.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

