Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
qplot(age, friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point()
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) +
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) +
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_jitter(alpha = 1/10) +
  xlim(13, 90)
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarize(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)


library(dplyr)
```

Create your plot!

```{r Conditional Means Plot}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
names(pf.fc_by_age)
ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y=friend_count), data = pf) +
  geom_point(alpha=0.05,
             position= position_jitter(h = 0),
             color = "orange") +
  coord_trans(y = 'sqrt') + 
  geom_line(stat = "summary", fun.y = mean)+
  geom_line(stat="summary", fun.y = quantile, probs = 0.1,
            linetype =2, color="blue") + 
  geom_line(stat="summary", fun.y = quantile, probs = 0.5,
            linetype =2, color="blue") + 
  geom_line(stat="summary", fun.y = quantile, probs = 0.9,
            linetype =2, color="blue")
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$friend_count, pf$age)

with(pf, cor.test(friend_count, age))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
?subset
with(subset(pf, age <= 70)                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x=www_likes_received, y=likes_received), data = pf) +
  geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x=www_likes_received, y=likes_received), data=pf) +
  geom_point() +
  xlim(0,quantile(pf$www_likes_received, 0.95)) +
  ylim(0,quantile(pf$likes_received, 0.95)) +
  geom_smooth(method="lm", color="red")
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received,pf$likes_received)
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data(Mitchell)
names(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month, y=Temp), data=Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=(Month%%12), y=Temp), data=Mitchell)+
  geom_point() 

ggplot(aes(x=Month, y=Temp), data=Mitchell) +
  geom_point() +
  scale_x_discrete(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + (1.0 - pf$dob_month/12)
names(pf)
head(pf$age_with_months)
```

***

### Age with Months Means

```{r Age with Months Means}
age_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_months_groups,
                                 friend_count_mean=mean(friend_count),
                                 friend_count_median=median(friend_count),
                                 n=n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}
ggplot(aes(y=friend_count_mean, x=age_with_months),
       data = subset(pf.fc_by_age_months, age_with_months <= 71)) +
  geom_line()
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(y=friend_count_mean,x=age_with_months),
       data=filter(pf.fc_by_age_months, age_with_months<=71))+
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(age,friend_count_mean), data =subset(pf.fc_by_age, age<=71)) +
  geom_line()+
  geom_smooth()


p2 <-  ggplot(aes(y=friend_count_mean,x=age_with_months),
       data=filter(pf.fc_by_age_months, age_with_months<=71))+
  geom_line()+
  geom_smooth()

p3 <- ggplot(aes(y= friend_count, x = round(age/5)*5),
             data=subset(pf, age<= 71))+
  geom_line(stat="summary", fun.y = mean)


library(gridExtra)
grid.arrange(p1,p2,p3,ncol=1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:
We learn the comparison two variables in our data. We learned how figure out correlation between 2 variables. We learn how do visualization for analyze two variables.
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

