---
title: "Problem set 4"
author: "Shalva Usubov shaliko@ezid.ru"
date: "5 апреля 2015 г."
output: html_document
---

Scatterplot of price vs x. using the ggplot syntax.
```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)
library("ggthemes")

data(diamonds)
names(diamonds)
```

```{r}
ggplot(aes(x=x, y=price), data=diamonds) + geom_point()
```

```{r, echo=FALSE}
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
```


scatter plot price vs depth
```{r}
ggplot(aes(x=depth, y=price), data=diamonds) + geom_point()
```

Adjustments - price vs. depth
```{r}
ggplot(aes(x=depth, y=price), data=diamonds) + 
  geom_point(alpha = 1/100) +
  scale_x_continuous(breaks = seq(50,70,2))
```

Correlation
```{r}
cor.test(diamonds$depth, diamonds$price)
```

Price vs. carat (omit the top 1% of price and carat)
```{r}
ggplot(aes(x=carat, y=price), data=diamonds) +
  geom_point(stat = "summary", fun.y = quantile, fun.x = quantile, probs = 0.99)
```

Price vs. volume
```{r}
diamonds$volume <- ((diamonds$x * diamonds$y) * diamonds$z)
ggplot(aes(x=volume, y=price), data=diamonds) +
  geom_point()
```

Observations:
Looks like we see linear correlation and that is highly correlation betwen volume and price. But we habe a some outliers? Some volumes are 0 and some value is very hight (~3800) but have "middle" price (~12500).

Correlations on Subsets
```{r}
subset_of_diamonds_by_volume <-subset(diamonds, volume>0 & volume <= 800) 
with(subset_of_diamonds_by_volume, cor.test(price, volume))
```

 Adjustments - price vs. volume
```{r}
subset_of_diamonds_by_volume <-subset(diamonds, volume>0 & volume <= 800) 
ggplot(aes(x=volume, y=price), data=subset_of_diamonds_by_volume)+
  geom_point(alpha=1/20) +
  geom_smooth(method='lm')
```

Mean Price by Clarity
```{r}
library(dplyr)

diamondsByClarity <- diamonds%>%
  group_by(clarity)%>%
  summarise(mean_price=mean(price),
            median_price=median(price),
            min_price=min(price),
            max_price=max(price),
            n=n())
head(diamondsByClarity)
```

Bar Charts of Mean Price
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(x=clarity, y=mean_price), data=diamonds_mp_by_clarity) +
  geom_bar(stat="identity")
p2 <- ggplot(aes(x=color, y=mean_price), data=diamonds_mp_by_color) +
  geom_bar(stat="identity")
grid.arrange(p1,p2,ncol=1)
```

Observation:
Mean price by clarity: The most expensive clarity by mean is "SL2".
Mean price by color: The most expensive color by mean is "J" and price increase in alphabetical order: "D" is lower and "J" is highest.


Gapminder Revisited
```{r}
bl <- read.csv("number_of_billionaires_total.csv", header = T, row.names=1)
total <- colSums(bl, na.rm=T)
average <- colMeans(bl, na.rm = T)
df_y <- as.data.frame(cbind(total, average))
df_y$Year <- as.numeric(substr(rownames(df_y), start=2, stop=5))
head(df_y)
p1 <- ggplot(aes(x=Year, y=total), data=df_y) + geom_point()
p2 <- ggplot(aes(x=Year, y=average), data=df_y) + geom_point()
grid.arrange(p1,p2,ncol=1)
```

```{r}
bl <- read.csv("number_of_billionaires_total.csv", header = T, row.names=1)
average <- colMeans(bl, na.rm = T)
df_y <- as.data.frame(average)
df_y$Year <- as.numeric(substr(rownames(df_y), start=2, stop=5))
ggplot(aes(x=Year, y=average), data=df_y) + 
  geom_point() +
  geom_smooth(method='lm', stat="identity") +
  theme_hc() +
  ggtitle('Average/Year of billionaires in the world') +
  ggsave('avgbilplot.jpg')
```
