```{r}
library(ggplot2)
data(diamonds)

ggplot(diamonds, aes(x=price)) + geom_histogram()


summary(diamonds$price)

nrow(diamonds[diamonds$price < 500,])
nrow(diamonds[diamonds$price < 250,])
nrow(diamonds[diamonds$price >= 15000,])

ggplot(aes(x=price), data = diamonds) + geom_histogram(binwidth = 250)+
  scale_x_continuous(breaks = seq(0,5000,250), limit = c(0,5000))

#For diamonds that cost less than $2,000, the most common price of a diamond is around $700 (binwidth = 1)

ggplot(aes(x=price), data = diamonds) + geom_histogram(binwidth = 250)+
  scale_x_continuous(breaks = seq(0,5000,250), limit = c(0,5000)) + facet_wrap(~cut, ncol=2)


by(diamonds$price,diamonds$cut,summary)
by(diamonds$price,diamonds$cut,max)


qplot(x = price, data = diamonds) + facet_wrap(~cut)

qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = "free_y")

ggplot(aes(x=price/carat), data = na.omit(diamonds)) + geom_histogram()+
  facet_wrap(~cut, scales  = "free_y") + scale_x_log10()


ggplot(aes(x = color, y = price/carat), data = na.omit(diamonds)) + geom_boxplot()
```

#Interquartile Range - IQR
```{r}
summary(subset(diamonds,color=='D')$price)
summary(subset(diamonds,color=='J')$price)
IQR(subset(diamonds, color == 'D')$price)
by(diamonds$price, diamonds$color, IQR)
```


#Price per Carat Box Plots by Color
```{r}
ggplot(aes(x = color, y = price/carat), data = diamonds) + geom_boxplot(aes(fill=color)) +
  ggsave('pricePerCaratInColor.png')
```

#Carat Frequency Polygon
```{r}
table(diamonds$carat)[table(diamonds$carat)>2000]
```

#Gapminder Data
```{r}
al = read.csv('alcohol_indicator.csv', header=TRUE, row.names=1)
ggplot(data = al) + 
  geom_freqpoly(aes(color = 'Year', x = al$X2005, y = (..count..)/sum(..count..))) + 
  geom_freqpoly(aes(color = 'blue', x = al$X2008, y = (..count..)/sum(..count..))) +
  xlab('Сount litre per people in a year') +
  ylab('Average drink over countries') +
  scale_colour_manual(values = c('green','blue'),
                      labels = c('2005','2008')) +
  ggsave('al.png')
```

#Exploring Your Friends' Birthdays
```{r}
library(lubridate)
library(gridExtra)

fb = read.csv('facebook_friends.csv', stringsAsFactors=FALSE)
head(fb$Birthday)

fb$b_date <- mdy(fb$Birthday)
fb$b_year <- year(fb$b_date)
fb$b_month <- month(fb$b_date)
fb$b_day <- day(fb$b_date)
```
## How many people share your birthday? Do you know them?
```{r}
NROW(fb$Birthday)
```
## Which month contains the most number of birthdays?
```{r}
tbmonth = table(fb$b_month)

which(tbmonth == max(tbmonth))
```
## How many birthdays are in each month?
```{r}
table(fb$b_month)
```
## Which day of the year has the most number of birthdays?
```{r}
tbday <- table(fb$b_day)

which(tbday == max(tbday))
```
## Do you have at least 365 friends that have birthdays on everyday of the year?
```{r}             
if ((NROW(fb$b_day) > 365) == F) {
  paste(c("No, I have only", as.character(NROW(fb$b_day)), 'friends'), collapse= " ")
}
```